---
title: A simple rmarkdown example
subtitle: Open Online Introduction to R Course
author: |
  | Wolfgang Viechtbauer
  | Maastricht University
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  # bookdown::html_document2:
  html_document:
    toc: true
    number_sections: false
    toc_depth: 2
    toc_float:
      collapsed: false
    theme: default # lots of nice themes can be used: https://bootswatch.com/
    highlight: haddock # default, tango, pygments, kate, monochrome, espresso, zenburn, haddock
    df_print: default
  # pdf_document:
  # # bookdown::pdf_book:
  #   toc: true
  #   number_sections: false
  #   toc_depth: 2
  # word_document
fig_caption: yes
bibliography: references.bib
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# code to install/load packages (at the very beginning of the document)

if (!suppressWarnings(require(pander))) install.packages("pander")
library(pander)

if (!suppressWarnings(require(knitr))) install.packages("knitr")
library(knitr)

if (!suppressWarnings(require(texreg))) install.packages("texreg")
library(texreg)
```

# Introduction

This is an example of a R markdown document. With some basic formating codes you can make text *italic* and **bold**.

## Subheading

You can also easily make lists:

* item 1
* item 2
* item 3

Or numbered lists:

1. item 1
2. item 2
3. item 3

### Subsubheading

This is a "subsubheading". [^1]

# Equations

If you know some basic latex syntax, you can easily add equations to your document. Equations can appear as part of the regular text (e.g., if $x = 2$, then $x + 4 = 6$). Or you can have equations appear in their own line: $$\frac{1}{2} + \frac{3}{4} = \frac{5}{4}.$$ When you compile the document, the equations should be nicely rendered.

Here is a more interesting example. The probability density function of a normal distribution is given by $$f(x) = \frac{1}{\sqrt{2\pi} \sigma} e^{-\frac{1}{2}\left(\frac{x-\mu}{\sigma}\right)^2},$$ where $\mu$ denotes the mean of the distribution and $\sigma$ its standard deviation.

# Links

Links can be created by simply pasting the URL into the document: https://rmarkdown.rstudio.com/. Or, you can make a string of text become a link. For example: To learn more, go to the [R markdown](https://rmarkdown.rstudio.com/) website.

# Embedding R Code

Where it gets really cool is when you start embedding R code in your document. For example:

```{r}
set.seed(1234)
x <- runif(100, min=0, max=30)
y <- 10 + .5*x + .3*x^2 + rnorm(100, mean=0, sd=25)
res1 <- lm(y ~ x)
res2 <- lm(y ~ x + I(x^2))
summary(res1)
```

Of course, in a manuscript or presentation, you may not want to show the code itself, just the output. That is of course possible by using `echo=FALSE` (possibly combined with `message=FALSE` and/or `warning=FALSE` as needed when starting a "code chunk").

## Dynamically Generated Plots

Using regular R code, you can also create graphs, which then get automatically embedded in the document.

```{r models, fig.width=7, fig.height=5.5, fig.align='center', fig.cap='R Skills as a Function of Time in the Course'}
par(mar=c(5,4,2,2))
plot(x, y, pch=19, xlab="Time in R Course", ylab="Your R Skills")
abline(res1, lty="dotted")
lines(predict(res2, newdata=data.frame(x=seq(0,30,1))))
legend("topleft", inset=.02, legend=c("Linear Model", "Quadratic Model"), lty=c("solid", "dotted"))
```

You can also include the results from analyses in your text. For example, for the quadratic model, we find $R^2 =$ `r round(summary(res2)$r.squared, 2)`. I think we should send this article to *Science* or *Nature*! When using `bookdown::html_document2` as the output format, figures are automatically numbered and you can refer to them from the text. For example, see Figure \@ref(fig:models) (note that this won't work when using `html_document` as the output format).

## Tables

You can also create tables from model objects (the [pander](https://cran.r-project.org/package=pander) package is very useful for that).

```{r, echo=FALSE, results='asis'}
pander(res2, caption="Results from Quadratic Model")
```

The `kable()` function from the [knitr](https://cran.r-project.org/package=knitr) package can do something similar.

```{r, echo=FALSE, results='asis'}
kable(round(coef(summary(res2)), 3), caption="Results from Quadratic Model")
```

The [gt](https://gt.rstudio.com/), [xtable](https://cran.r-project.org/package=xtable), [stargazer](https://cran.r-project.org/package=stargazer), and [flextable](https://cran.r-project.org/package=flextable) packages can also create pretty tables. I recently stumbled across the [texreg](https://cran.r-project.org/package=texreg) package, which can create nicely formatted tables based on one or multiple regression models.

```{r, echo=FALSE, results='asis'}
knitreg(list(res1, res2), caption="Comparisom of the Linear and Quadratic Models")
```

These are just some of the basics. I hope you can see the potential here!

# References

You can also include references. You need a bibliography file for this (see the `references.bib` file). For example, the first paper about the R language was written by @ihaka1996. Blah blah blah. We used R for the analyses [@r2020].

# More Info

To learn more about R Markdown, you may want to check out these websites:

* [rmarkdown](https://rmarkdown.rstudio.com/)
* [markdown](https://daringfireball.net/projects/markdown/) (a nice [tutorial](https://commonmark.org/help/tutorial/))
* [R Markdown: The Definitive Guide](https://bookdown.org/yihui/rmarkdown/)
* [bookdown: Authoring Books and Technical Documents with R Markdown](https://bookdown.org/yihui/bookdown/)
* [Advanced R Markdown](https://rpubs.com/bpbond/630335) by Ben Bond-Lamberty

***

[^1]: A footnote! The numbering is done automatically when the document is rendered, so you could also use `[^chicken]` to insert a new footnote.

# References
